### 2023-02-12 2023 Disambiguation/Disambiguation.tex
Article on disambiguation

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithmic}
\usepackage{bbold}
\usepackage[cal=boondoxo]{mathalfa}
\begin{document}

\title{What's in a name? Disambiguation strategies for relational descriptions}
\author{Sebastiaan Joosten}
\date{Work in progress}
\maketitle

\section{Motivation}
Notations matter: They help communicate ideas by establishing expectations.
When programming, it is recommended to choose function symbols, variable names, etcetera, in a way that it helps the reader of the program.
Notations refer to a certain semantical notion, and choosing a good notation can help establish that.
Regardless of expectations, the thing a notation refers to must always be clear.

Programming languages greatly influence the notations a programmer uses through the rules on what a notation might refer to.
This paper looks at how identifiers that refer to binary relations might be used in a semantic modeling language.
In particular, we look at the use of identifiers in the programming language Ampersand, and ways in which changes to the language might effect the use of identifiers.

Modeling in Ampersand happens via the use of binary relations, which are declared by stating a name, a source concept, and a target concept.
When referring to binary relations, we typically only refer to them by their name.
It is desirable to allow different binary relations to have the same name:
Allowing this gives the programmer greater flexibility in which names to use.

\subsection{Running example}
An application can be described by binary relations over concepts, a description of what the concepts and relations mean, stating invariants and process rules that determine the behavior of the system, and describing interfaces to interact with the system.
For the purposes of this article, we just declare binary relations and some invariants.

We give the following planning application as an example in which binary relations share a name:
\begin{verbatim}
before : Task * Task
before : Time * Time
scheduled : Task * Time
\end{verbatim}

An invariant in this system is:

\[
\mathrm{before}_{\mathrm{Task * Task}};\mathrm{scheduled}_{\mathrm{Task * Time}} \subseteq \mathrm{scheduled};\mathrm{before}_{\mathrm{Time * Time}}
\]

Such invariants are well-typed: a Task is only compared to a Task, and a Time is only compared to a Time, and the only way in which a Task and a Time are related is through explicitly declared relations.
We can consider the same expression in which types are dropped: $\mathrm{before};\mathrm{scheduled} \subseteq \mathrm{scheduled};\mathrm{before}$.
If we require that invariants are well-typed, there is only one way of annotating such expressions with types, regardless of our interpretation.

In Ampersand's implementation, Ampersand does not require writing the types: the use of the possibly ambiguous relation name $\mathrm{before}$ is allowed as long as it is clear which relation is intended, as is the case here.
We refer to the problem of determining which relation is intended as the disambiguation problem, which we investigate more closely in a separate section.

\subsection{Generalization}
A common task in semantic modeling is to identify that some concepts are generalizations or specializations of other concepts.
This section motivates the need for a special language feature to deal with generalizations.

As an example, our planning application allows regular working hours to be set, which means that some times fall during regular hours.
This means that such times would be a specialization of a time.
We can model this with a relation called $\mathrm{worktime}$ and an invariant: $\mathrm{worktime} \subseteq \mathbb{1}_{Time}$, where $\mathbb{1}_{A} = \{(a,a) \mid a\in A \}$ is the relational identity.
Another way to model this, is to introduce a concept $\mathrm{Worktime}$ and use $\mathbb{1}_\mathrm{Worktime}$ in the place of the relation $\mathrm{worktime}$, along with the invariant $\mathbb{1}_\mathrm{Worktime} \subseteq \mathbb{1}_{Time}$.
However, this latter invariant is ill-typed: the expression on the left of the subset ($\subseteq$) is of type $(\mathrm{Worktime},\mathrm{Worktime})$, while the expression on the right is of type $(\mathrm{Time},\mathrm{Time})$.
To allow expressions like $\mathbb{1}_\mathrm{Worktime} \subseteq \mathbb{1}_{Time}$ in a programming language, we need a more refined notion of which expressions are `type correct'.
A paper by van der Woude and Joosten shows how one can do this~\cite{Woude11}.
However, that paper does not consider the interplay with disambiguation which we consider here.

\subsection{Two interpretations.}
There are two different ways in which we can think of the two binary relations called $\mathrm{before}$.
In one interpretation, we can say that $\mathrm{before}$ is a single binary relation, defined over the (disjoint) union between Task and Time.
Whenever we use $\mathrm{before}$, we restrict its type to either \verb=Task * Task= or \verb=Time * Time= to get two completely separate behaviors.
This means that we read $\mathrm{before}_{\mathrm{Task * Task}}$ as $\mathrm{before} \cap (\mathrm{Task}\times\mathrm{Task})$.

Another interpretation is to say that there are two entirely separate relations, $\mathrm{before}_{\mathrm{Task * Task}}$ and $\mathrm{before}_{\mathrm{Time * Time}}$.
In this interpretation, we can consider ${\mathrm{Task * Task}}$ as a part of the name of the relation.

Note our interpretation does not matter for the behavior we get from well-typed invariants like the one we saw before.
However, a problem arises if we add generalization.

If we were to state - for some obscure reason - that every Task is a Time, in our first interpretation we would expect that $\mathrm{before}_{\mathrm{Task * Task}} = \mathrm{before}_{\mathrm{Time * Time}} \cap (\mathrm{Task}\times\mathrm{Task})$, since $(\mathrm{Task}\times\mathrm{Task}) \subseteq (\mathrm{Time}\times\mathrm{Time})$.
In our second interpretation, we would not expect the behavior to change.

Observe that we are free to add $\mathrm{before}_{\mathrm{Task * Task}} \subseteq \mathrm{before}_{\mathrm{Time * Time}}$, or the stronger statement $\mathrm{before}_{\mathrm{Task * Task}} = \mathrm{before}_{\mathrm{Time * Time}} \cap (\mathrm{Task}\times\mathrm{Task})$, as an invariant under the second interpretation.
Therefore, we conclude that working with the second interpretation is most flexible and propose to follow this interpretation.

At the time of writing, Ampersand infers the invariant $\mathrm{before}_{\mathrm{Task * Task}} \subseteq \mathrm{before}_{\mathrm{Time * Time}}$ without requiring that $\mathrm{before}_{\mathrm{Task * Task}} = \mathrm{before}_{\mathrm{Time * Time}} \cap (\mathrm{Task}\times\mathrm{Task})$. This can be seen as a sort of middle ground between the two interpretations. However, we foresee that Ampersand will change its implementation and do not follow its behavior here.

\section{Solving the disambiguation problem without generalization}
As an aside, we ask whether a compiler can be expected to figure out which relations are intended in reasonable time. In this section, we answer this question for a very simple setting in the affirmative by giving a simple efficient algorithm.

We can define the disambiguation problem as follows:

A disambiguation problem is given by an expression $E$ using the operations $\cap$ and $;$, and relation names from a set of names $\mathcal{N}$, and a set of possible types of each name: $\mathcal{t}(n) \subseteq \mathcal{T}\times \mathcal{T}$ for $n\in \mathcal{N}$.
A solution to the disambiguation problem is a well-typed expression $S$ in which the annotations to relation names corresponds to an element in $\mathcal{t}(n)$, such that dropping the types from the expression yields $E$.

As an example, $E = (\mathrm{before};\mathrm{scheduled}) \cap (\mathrm{scheduled};\mathrm{before})$ with types $\mathcal{t}(\mathrm{before}) = \{(\mathrm{Task}, \mathrm{Task}),(\mathrm{Time}, \mathrm{Time})\}$ and $\mathcal{t}(\mathrm{scheduled}) = \{(\mathrm{Task}, \mathrm{Time})\}$, is a disambiguation problem.
A solution would be:
\[\left(\mathrm{before}_{\mathrm{Task * Task}};\mathrm{scheduled}_{\mathrm{Task * Time}}\right) \cap \left(\mathrm{scheduled};\mathrm{before}_{\mathrm{Time * Time}}\right)\]

We can solve the disambiguation problem using two passes through $E$.
In the first pass, we decorate the expression tree for $E$ with all the possible types for each sub-expression bottom up.
Let $p(e)$ denote the possible types, then $p(n) = \mathcal{t}(n)$, and $p(e_1;e_2) = \{(x,y) \mid \exists z. (x,z) \in p(e_1) \wedge (z,y) \in p(e_2))\}$, and $p(e_1 \cap e_2) = p(e_1) \cap p(e_2)$.
If $p(E)$ is the empty set, then there is no solution to the disambiguation problem, and we are done.
If $p(E)$ contains multiple elements, then there are multiple solutions to the disambiguation problem.
We can arbitrarily select a single element from $p(E)$ to get an arbitrary solution\footnote{In practice, we might want to indicate that $E$ is ambiguous, which could require running the next step twice: once for two different elements in $p(E)$.}.

For the top-down step, let $d(e,t)$ denote the disambiguated expression $e$ with the type $t$.
For the top-level call of $d(e,t)$, let $e=E$ and $t\in p(E)$.
Then $d(n,t) = n_{[t]}$, $d(e_1 \cap e_2, t) = d(e_1,t) \cap d(e_2,t)$ and $d(e_1 ; e_2, (x,y)) = d(e_1, (x,z)) ; d(e_2, (z,y))$ where $z$ is picked such that $(x,z)\in p(e_1)$ and $(z,y)\in p(e_2)$.
Such a $z$ is guaranteed to exist by how $p$ is defined.

As an example, let's look at $E = (\mathrm{before};\mathrm{scheduled}) \cap (\mathrm{scheduled};\mathrm{before})$.
Then $p(\mathrm{before}) =  \{(\mathrm{Task}, \mathrm{Task}),(\mathrm{Time}, \mathrm{Time})\}$ following $\mathcal{t}$ and similarly $p(\mathrm{scheduled}) = \{(\mathrm{Task}, \mathrm{Time})\}$.
This gives subterms $p(\mathrm{before};\mathrm{scheduled}) = p(\mathrm{scheduled};\mathrm{before}) = \{(\mathrm{Task}, \mathrm{Time})\}$ so $p(E) = \{(\mathrm{Task}, \mathrm{Time})\}$.

Top-down, $d(E,t) = d(\mathrm{before};\mathrm{scheduled},t) \cap d(\mathrm{scheduled};\mathrm{before},t)$ with $t=(\mathrm{Task}, \mathrm{Time})$.
Since $E$ is not ambiguous, we have only one possible way to continue as we solve the disambiguation problem:
\begin{align*}d(\mathrm{before};\mathrm{scheduled},t) &= d(\mathrm{before},(\mathrm{Task}, \mathrm{Task})) ;  d(\mathrm{scheduled},(\mathrm{Task}, \mathrm{Time})) \\
&=\mathrm{before}_{\mathrm{Task} * \mathrm{Task}}; \mathrm{scheduled}_{\mathrm{Task}* \mathrm{Time}}\\
d(\mathrm{scheduled};\mathrm{before},t) &=  d(\mathrm{scheduled},(\mathrm{Task}, \mathrm{Time}));d(\mathrm{before},(\mathrm{Time}, \mathrm{Time}))\\
&=\mathrm{scheduled}_{\mathrm{Time} * \mathrm{Task}}; \mathrm{before}_{\mathrm{Time}* \mathrm{Time}}\\
\end{align*}

The disambiguation problem and the corresponding algorithm can be extended over all operations in relation algebra.

\section{Interplay between Generalization and Disambiguation}

% van der Woude paper: https://link.springer.com/chapter/10.1007/978-3-642-21070-9_25
\end{document}